---
title: "Tyre Wear emission model analysis"
output: md_document
editor_options: 
  chunk_output_type: inline
---

```{r prepare, include=FALSE}
library(tidyverse)
```

Tyre wear model by Joris Meesters and Joris Quik

## Primary algorithms

```{r primary algorithms}
#### Friction force functions ####
# Roll resistance force in N
f_roll_force <- function(c_roll, m_vehicle, grav_constant) {c_roll*m_vehicle*grav_constant}

# Drag resistance force in N
f_drag_force <- function(c_drag, A_vehicle, rho_air, v_vehicle, v_wind) {pmax(0,c_drag*A_vehicle*rho_air*(((v_vehicle)+v_wind)^2)/2)}

# Slope force
f_slope_force <- function(m_vehicle, grav_constant, alpha_slope) {m_vehicle*grav_constant*sin(alpha_slope)}

# Brake force
f_brake_force <- function(m_vehicle, m_rotate, c_brake) {(m_vehicle+m_rotate)*c_brake}

# Inertia force upon acceleration
f_accel_force <- function(m_vehicle, m_rotate, c_accel) {(m_vehicle+m_rotate)*c_accel}

# Inertia force upon deceleration (laying foot of gas)
f_c_decel_inert <- function(Decel_force,m_vehicle,m_rotate){Decel_force/(m_vehicle+m_rotate)}

# Centripetal force in cornering
f_centripet_force <- function(m_vehicle , v_vehicle , r_corner) {m_vehicle*(v_vehicle^2)/r_corner}

# Bank force in cornering
f_bank_force <-function(grav_constant, alpha_bank_slope, m_vehicle){grav_constant*sin(alpha_bank_slope)*m_vehicle}

#### Distance functions ####

# Acceleration distance in m
# function of acceleration time
f_accel_time<-function(v_start,v_end,c_accel){-(v_start-v_end)/c_accel}

# function acceleration distance
f_accel_distance <- function(v_start , Accel_time , c_accel ) {v_start*Accel_time+1/2*c_accel*Accel_time^2}

# Deceleration distance in m
f_decel_time<-function(v_start,v_end,c_decel){(v_start-v_end)/c_decel}

f_decel_distance <- function(v_start , Decel_time , c_decel ) {v_start*Decel_time+1/2*c_decel*Decel_time^2}

f_longitude_friction_work <- function(long_force , slip, distance){(long_force)*slip*distance}
f_latitude_friction_work <- function(lat_force, slip, distance){lat_force*slip*distance}

# Corner distance in m
f_corner_distance <- function(r_corner,corner_angle){(corner_angle/(360)*2*r_corner*pi)}

### Slip Functions ###
f_normal_load_force <- function(m_vehicle, grav_constant){m_vehicle*grav_constant}
f_mu_slip <- function(normal_load_force, long_force){normal_load_force/long_force}
f_mu_max <- function(grip_index_tyre, x_correct_road, mu_max_ref_tyre){(grip_index_tyre)/((1/mu_max_ref_tyre)*(125/100))*x_correct_road}
f_x_slip_long_force <- function(mu_max, normal_load_force, optimal_slip){(optimal_slip)/(mu_max*normal_load_force)}
f_slip_wheelspin <- function (optimal_slip, x_slip_long_force, long_force){pmin(optimal_slip, x_slip_long_force*long_force)} 

f_c_full_brake <-function(grip_index_tyre,
                          x_correct_road,
                          c_full_brake_ref_tyre){
  (grip_index_tyre)/((1/c_full_brake_ref_tyre)*(125/100))*x_correct_road
}

f_slip_brake <- function (c_decel_brake, c_full_brake){c_decel_brake/c_full_brake}
f_mu_lat_slip <- function(normal_load_force, long_force){normal_load_force/lat_force} 
f_slip_lateral <- function (optimal_slip, x_slip_lat_force, lat_force){pmin(optimal_slip, x_slip_lat_force*lat_force)}



```

## Maneuver Functions

### acceleration dry

```{r functions for different maneuvers}
## Simulating longitudinal friction work per track sector under dry conditions
### dry condition slip parameters

f_acc_dry <- function(sector_velocity_ms,
                      sector_start_velocity_ms,
                      frac_driver_comfort_max_accelaration,
                      c_accel_max,
                      m_vehicle,
                      m_rotate,
                      c_drag,
                      A_vehicle,
                      rho_air,
                      v_wind,
                      c_roll,
                      grav_constant,
                      sector_alpha_slope,
                      optimal_slip_dry_asphalt,
                      x_slip_long_force_dry_asphalt,
                      accel_force_end_maneuver_1 = 0){
  
  
  maneuver_1 = "acceleration on dry asphalt"
  v_start_acceleration_maneuver_1 = pmin(sector_velocity_ms,sector_start_velocity_ms) 
  
  
  v_end_acceleration_maneuver_1 = sector_velocity_ms
  v_average_maneuver_1 = (v_start_acceleration_maneuver_1 + v_end_acceleration_maneuver_1)/2
  
  c_accel_maneuver_1 <- frac_driver_comfort_max_accelaration * c_accel_max
  
  accel_force_t0_maneuver_1 = f_accel_force(m_vehicle = m_vehicle, 
                                            m_rotate = m_rotate, 
                                            c_accel = c_accel_maneuver_1)
  
  
  accel_force_average_maneuver_1 = (accel_force_t0_maneuver_1 + accel_force_end_maneuver_1)/2
  
  duration_maneuver_1 = f_accel_time(v_start = v_start_acceleration_maneuver_1, 
                                     v_end = v_end_acceleration_maneuver_1, 
                                     c_accel = c_accel_maneuver_1)
  distance_maneuver_1 = f_accel_distance(v_start = v_start_acceleration_maneuver_1, 
                                         Accel_time = duration_maneuver_1, 
                                         c_accel = c_accel_maneuver_1)
  
  drag_force_maneuver_1 = f_drag_force(c_drag = c_drag, 
                                       A_vehicle = A_vehicle, 
                                       rho_air = rho_air, 
                                       v_vehicle = v_average_maneuver_1, 
                                       v_wind = v_wind)
  roll_force_maneuver_1 = f_roll_force(c_roll=c_roll , 
                                       m_vehicle = m_vehicle, 
                                       grav_constant = grav_constant) 
  
  # so here you implement to take 0 if the slope is going downhill for the force related to the slope. 
  # What about the reduced need to accelerate under a negative slope? This does not seem to be implemented.
  slope_force_maneuver_1 = pmax(0,f_slope_force(m_vehicle = m_vehicle, 
                                                grav_constant = grav_constant, 
                                                alpha_slope = sector_alpha_slope))
  
  slip_maneuver_1 = f_slip_wheelspin(optimal_slip = optimal_slip_dry_asphalt, 
                                     x_slip_long_force = x_slip_long_force_dry_asphalt, 
                                     long_force = (accel_force_average_maneuver_1+ 
                                                     drag_force_maneuver_1+ 
                                                     roll_force_maneuver_1+ 
                                                     slope_force_maneuver_1)) # changed as pmin(0,slope_force_maneuver_1) is not expected and probably incorrect as it is always 0
  
  longitude_friction_work_maneuver_1 = f_longitude_friction_work(long_force = (drag_force_maneuver_1 + roll_force_maneuver_1 + slope_force_maneuver_1 + accel_force_average_maneuver_1), slip = slip_maneuver_1, distance = distance_maneuver_1)
  
  # this implementation results in relationship between distance of the this maneuver and the required time needed to go from starting velocity to end velocity. Care should be taken to then allow the rest of the track distance to be driven at constant speed.
  # for this reason a list is reported of both
  list(maneuver=maneuver_1,distance=distance_maneuver_1, friction_work_longitude = longitude_friction_work_maneuver_1)
}

```

### deceleration dry

```{r functions for different maneuvers}
f_dec_dry <- function(sector_velocity_ms,
                      sector_end_velocity_ms,
                      c_decel_maneuver_2 = c_decel_comfort,
                      m_vehicle,
                      m_rotate,
                      c_drag,
                      A_vehicle,
                      rho_air,
                      v_wind,
                      c_roll,
                      grav_constant,
                      sector_alpha_slope,
                      c_full_brake_dry_asphalt,
                      optimal_slip_dry_asphalt,
                      x_slip_long_force_dry_asphalt){
  maneuver_2 = "deceleration on dry asphalt"
  
  v_start_deceleration_maneuver_2 = sector_velocity_ms
  v_end_deceleration_maneuver_2 = pmin(sector_velocity_ms, 
                                       sector_end_velocity_ms) # Can these functions be merged?
  v_average_maneuver_2 = (v_start_deceleration_maneuver_2 + 
                            v_end_deceleration_maneuver_2)/2
  drag_force_t0_maneuver_2 = f_drag_force(c_drag, A_vehicle, 
                                          rho_air, 
                                          v_vehicle = v_start_deceleration_maneuver_2, 
                                          v_wind)
  drag_force_average_maneuver_2 = f_drag_force(c_drag, 
                                               A_vehicle, 
                                               rho_air, 
                                               v_vehicle = v_average_maneuver_2, 
                                               v_wind)
  roll_force_maneuver_2 = f_roll_force(c_roll, 
                                       m_vehicle, 
                                       grav_constant)
  slope_force_maneuver_2 = f_slope_force(m_vehicle = m_vehicle, 
                                         grav_constant = grav_constant, 
                                         alpha_slope = sector_alpha_slope)
  inert_decel_force_t0_maneuver_2 = (drag_force_t0_maneuver_2 + 
                                       roll_force_maneuver_2 + 
                                       pmax(0,slope_force_maneuver_2)) # What about negative slopes? HOw is that dealt with?
  c_decel_inert_t0_maneuver_2 = f_c_decel_inert(Decel_force = inert_decel_force_t0_maneuver_2, 
                                                m_vehicle = m_vehicle, 
                                                m_rotate = m_rotate) # waarom wordt dit uitgerekend (nergens gebruikt?)
  
  ## Brake force needed to descell comfortable
  
  brake_force_maneuver_2 = pmax(0,
                                (m_vehicle + m_rotate) * c_decel_maneuver_2 - 
                                  drag_force_average_maneuver_2 - 
                                  roll_force_maneuver_2 - 
                                  slope_force_maneuver_2)
  
  c_comfort_brake_maneuver_2 = brake_force_maneuver_2 * 1/(m_vehicle + m_rotate)
  
  slip_brake_maneuver_2 = f_slip_brake(c_decel_brake =c_comfort_brake_maneuver_2, c_full_brake = c_full_brake_dry_asphalt)
  
  slip_maneuver_2= slip_brake_maneuver_2 + f_slip_wheelspin(optimal_slip = optimal_slip_dry_asphalt, 
                                                            x_slip_long_force = x_slip_long_force_dry_asphalt, 
                                                            long_force = drag_force_average_maneuver_2 + 
                                                              roll_force_maneuver_2 + 
                                                              pmin(0,slope_force_maneuver_2))
  
  
  duration_maneuver_2 = f_decel_time(v_start = v_start_deceleration_maneuver_2, 
                                     v_end = v_end_deceleration_maneuver_2, 
                                     c_decel = c_decel_maneuver_2)
  distance_maneuver_2 = f_decel_distance(v_start = v_start_deceleration_maneuver_2, 
                                         Decel_time = duration_maneuver_2 , 
                                         c_decel = c_decel_maneuver_2)
  
  longitude_friction_work_maneuver_2 = f_longitude_friction_work(long_force = (drag_force_average_maneuver_2 + 
                                                                                 roll_force_maneuver_2 + 
                                                                                 pmax(0,slope_force_maneuver_2) + 
                                                                                 brake_force_maneuver_2), 
                                                                 slip = slip_maneuver_2, 
                                                                 distance = distance_maneuver_2)                    
  
  
  list(maneuver=maneuver_2,distance=distance_maneuver_2, friction_work_longitude=longitude_friction_work_maneuver_2)
  
}
```

### constant speed dry

```{r functions for different maneuvers}
f_const_dry <- function(sector_velocity_ms,
                        sector_end_velocity_ms,
                        m_vehicle,
                        m_rotate,
                        c_drag,
                        A_vehicle,
                        rho_air,
                        v_wind,
                        c_roll,
                        grav_constant,
                        sector_alpha_slope,
                        c_full_brake_dry_asphalt,
                        optimal_slip_dry_asphalt,
                        x_slip_long_force_dry_asphalt,
                        distance_maneuver_3 = pmax(0,sector_distance - distance_maneuver_1 - distance_maneuver_2)){
  # Longitudinal friction work during constant speed driving
  maneuver_3 = "constant speed driving"
  ## Constant speed driving is the maneuver at which the vehicle is driving over the track sector with a constant velocity.
  ##The longitudinal resistive forces during this maneuver are 
  ## the roll force
  roll_force_maneuver_3 = f_roll_force(c_roll, 
                                       m_vehicle, 
                                       grav_constant)
  ## the drag force 
  drag_force_maneuver_3 = f_drag_force(c_drag, 
                                       A_vehicle, 
                                       rho_air, 
                                       v_vehicle = sector_velocity_ms, 
                                       v_wind)
  ## the slope force (in case of uphill driving)
  slope_force_maneuver_3 = f_slope_force(m_vehicle = m_vehicle, 
                                         grav_constant = grav_constant, 
                                         alpha_slope = sector_alpha_slope)
  ## the brake force (in case downhill driving downhill using the brakes remain under the speed limit)
  brake_force_maneuver_3 = pmax(0,-(drag_force_maneuver_3 + roll_force_maneuver_3 +slope_force_maneuver_3)) 
  
  ## Slip during constant speed driving on dry asphalt
  c_decel_brake_maneuver_3 = pmin(c_full_brake_dry_asphalt , 
                                  f_c_decel_inert(Decel_force = brake_force_maneuver_3, 
                                                  m_vehicle = m_vehicle, 
                                                  m_rotate = m_rotate))
  slip_brake_maneuver_3 = f_slip_brake(c_decel_brake = c_decel_brake_maneuver_3, 
                                       c_full_brake = c_full_brake_dry_asphalt)
  slip_maneuver_3 = pmin(1,
                         f_slip_wheelspin(optimal_slip = optimal_slip_dry_asphalt, 
                                          x_slip_long_force = x_slip_long_force_dry_asphalt, 
                                          long_force = drag_force_maneuver_3 + 
                                            roll_force_maneuver_3 + 
                                            pmin(0,slope_force_maneuver_3)) + 
                           slip_brake_maneuver_3)
  
  
  
  duration_maneuver_3 = distance_maneuver_3*sector_velocity_ms
  longitude_friction_work_maneuver_3 = f_longitude_friction_work(long_force = (drag_force_maneuver_3 + 
                                                                                 roll_force_maneuver_3 + 
                                                                                 pmax(0,slope_force_maneuver_3) + 
                                                                                 brake_force_maneuver_3), 
                                                                 slip = slip_maneuver_3, 
                                                                 distance = distance_maneuver_3)
  list(maneuver=maneuver_3,
       distance=distance_maneuver_3,
       friction_work_longitude=longitude_friction_work_maneuver_3) 
}
```

### cornering dry

```{r functions for different maneuvers}

f_corn_dry <- function(sector_velocity_ms,
                       sector_corner_radius,
                       m_vehicle,
                       m_rotate,
                       sector_bank_slope,
                       sector_corner_angle,
                       c_drag,
                       A_vehicle,
                       rho_air,
                       v_wind,
                       c_roll,
                       sector_alpha_slope,
                       grav_constant,
                       c_full_brake_dry_asphalt,
                       optimal_slip_dry_asphalt,
                       x_slip_long_force_dry_asphalt){  
  # latitudinal friction work during cornering
  maneuver_4 = "cornering"
  centripet_force_maneuver_4 = f_centripet_force(m_vehicle = m_vehicle, 
                                                 v_vehicle = sector_velocity_ms, 
                                                 r_corner = sector_corner_radius)
  bank_force_maneuver_4 = f_bank_force(grav_constant = grav_constant,
                                       alpha_bank_slope = sector_bank_slope, 
                                       m_vehicle = m_vehicle)
  slip_lateral_maneuver_4 = f_slip_lateral(optimal_slip = optimal_slip_dry_asphalt, 
                                           x_slip_lat_force = 7*x_slip_long_force_dry_asphalt, 
                                           lat_force = centripet_force_maneuver_4 + 
                                             bank_force_maneuver_4)
  distance_maneuver_4 = f_corner_distance(corner_angle = sector_corner_angle , 
                                          r_corner = sector_corner_radius)
  latitude_friction_work_maneuver_4 = f_latitude_friction_work(lat_force = centripet_force_maneuver_4 + 
                                                                 bank_force_maneuver_4, 
                                                               slip = slip_lateral_maneuver_4, 
                                                               distance = distance_maneuver_4)
  
  
  drag_force_maneuver_3 = f_drag_force(c_drag, 
                                       A_vehicle, 
                                       rho_air, 
                                       v_vehicle = sector_velocity_ms, 
                                       v_wind)
  roll_force_maneuver_3 = f_roll_force(c_roll, 
                                       m_vehicle, 
                                       grav_constant)
  ## the slope force (in case of uphill driving)
  slope_force_maneuver_3 = f_slope_force(m_vehicle = m_vehicle, 
                                         grav_constant = grav_constant, 
                                         alpha_slope = sector_alpha_slope)
  ## the brake force (in case downhill driving downhill using the brakes remain under the speed limit)
  brake_force_maneuver_3 = pmax(0,-(drag_force_maneuver_3 + roll_force_maneuver_3 +slope_force_maneuver_3)) 
  ## Slip during constant speed driving on dry asphalt
  c_decel_brake_maneuver_3 = pmin(c_full_brake_dry_asphalt , 
                                  f_c_decel_inert(Decel_force = brake_force_maneuver_3, 
                                                  m_vehicle = m_vehicle, 
                                                  m_rotate = m_rotate))
  slip_brake_maneuver_3 = f_slip_brake(c_decel_brake = c_decel_brake_maneuver_3, 
                                       c_full_brake = c_full_brake_dry_asphalt)
  
  slip_maneuver_3 = pmin(1,
                         f_slip_wheelspin(optimal_slip = optimal_slip_dry_asphalt, 
                                          x_slip_long_force = x_slip_long_force_dry_asphalt, 
                                          long_force = drag_force_maneuver_3 + 
                                            roll_force_maneuver_3 + 
                                            pmin(0,slope_force_maneuver_3)) + 
                           slip_brake_maneuver_3)
  
  # longitude friction work during cornering
  longitude_friction_work_maneuver_4 = f_longitude_friction_work(long_force = (drag_force_maneuver_3 + 
                                                                                 roll_force_maneuver_3 + 
                                                                                 pmax(0,slope_force_maneuver_3) + 
                                                                                 brake_force_maneuver_3), 
                                                                 slip = slip_maneuver_3, 
                                                                 distance = distance_maneuver_4)
  
  list(maneuver=maneuver_4,
       distance=distance_maneuver_4,
       friction_work_longitude=longitude_friction_work_maneuver_4,
       friction_work_latitude = latitude_friction_work_maneuver_4) 
  
}


```

## Sensitivity analysis

### Data inputs

```{r data}
#### Overview of default data values #####
n_runs <- 10000
# input variables as used in calculations/sensitivity analysis

input_data <- tibble(
  #Gravitational constant in m.s^-2
  'grav_constant'=9.81,
  ##### Landscape data #####
  #Density of air in kg.m^-3
  'rho_air'=1.205,
  # Wind speed in m.s^-1
  'v_wind' = runif(n=n_runs ,-5,5),
  
  # Landscape slope in %
  'grade_slope' = runif(n=n_runs,-10,10)/100,
  'grade_bank_slope' = runif(n=n_runs,-10,10)/100,
  
  ##### Track-tire default data values #####
  #Roll resistance coefficient in kg/kg at dry asphalt
  'c_roll'= runif(n=n_runs,1,10)/1000,
  #Grip index tyre
  'grip_index_tyre' = runif(n=n_runs,1.09,1.55),
  # Peak friction coefficient of reference tyre in reference grip testing
  'mu_max_ref_tyre' = 0.85,
  # correction mu max for road underground
  'x_correct_road_wet_asphalt' = 1,
  'x_correct_road_dry_asphalt' =runif(n=n_runs,1.07,1.47),
  # optimal slip ratios per underground
  'optimal_slip_dry_asphalt' = runif(n=n_runs,0.15,0.20),
  'optimal_slip_wet_asphalt' = runif(n=n_runs,0.08,0.12),
  
  ##### Vehicle data #####
  #Vehicle acceleration from 0-100 km.h^-1 in s
  't_0_100kmh' =runif(n=n_runs,8,20),
  #Vehicle mass in kg
  'm_vehicle'=runif(n=n_runs,900,3000),
  #Vehicle surface area in m^2
  'A_vehicle'=runif(n=n_runs,2,4),
  #Vehicle aerodynamic drag coefficient
  'c_drag'=runif(n=n_runs,0.02,0.5),
  #vehicle turning diameter
  'd_turn_vehicle' = 9.8,
  
  ##### Driving style data #####
  # the percentage of the maximum acceleration constant at which the driver feels comfortable 
  # for uniform distribution
  'p_min_acc' = 10, # minimum % of the maximum acceleration constant at which the driver feels comfortable
  'p_max_acc' = 100, # maximum % of the maximum acceleration constant at which the driver feels comfortable
  
  #Comfortable deceleration constant
  'c_decel_comfort' = runif(n=n_runs,2,3),
  
  ##### Road data #####
  #Corner radius in m
  'r_corner' = runif(n=n_runs,10,50),
  # corner angle in degrees
  'angle_corner' = runif(n=n_runs,0,360)
)

# RSTUDIOn shortcut for forward pipe operator: Ctrl + Shift + M
# https://shiny.rivm.nl/kasstvdj/R_cursussen/Tidy-R_cursus/introductie-tidyverse.html#forward-pipe-operator

input_data <- input_data |> 
  mutate(
    # Landscape slope in Degrees
    'alpha_slope' = atan(grade_slope)*(180/pi),
    'alpha_bank_slope' = atan(grade_bank_slope)*(180/pi),
    
    #Brake deceleration constant of reference tyre in reference wet grip testing in m.s^-2
    'c_full_brake_ref_tyre' = 0.68*grav_constant,
    
    #Maximum acceleration constant in m.s^-2
    'c_accel_max' = (100*(1000/3600))/t_0_100kmh,
    
    # Total mass of rotating vehicle parts in kg
    'm_rotate'= (runif(n=n_runs, 13,15)/100)*m_vehicle,
    
    'frac_driver_comfort_max_accelaration' = runif(n=n_runs,p_min_acc,p_max_acc)/100
    
  )



```

Currently the functions depend on track data, which is prepared here.

```{r track data}
#### Friction forces per maneuver ####
##### Constant speed straight driving #####

# Scenario for urban traffic
kph_min <- 0 # kmph
kph_max <- 70 # kmph
# what is representative minimum radius for highway corner?
# What is representative corner angle? (max)
# what is representative slope min/max?

# Scenario for highway traffic
# kph_min <- 60 # kmph
# kph_max <- 140 # kmph
# what is representative minimum radius for highway corner?
# What is representative corner angle? (max)
# what is representative slope min/max?

# Scenario for stop and go traffic
# kph_min <- 60 # kmph
# kph_max <- 140 # kmph
# what is representative minimum radius for highway corner?
# What is representative corner angle? (max)
# what is representative slope min/max?

# Scenario for rural traffic
# kph_min <- 0 # kmph
# kph_max <- 90 # kmph
# what is representative minimum radius for highway corner?
# What is representative corner angle? (max)
# what is representative slope min/max?


input_data <- input_data |> mutate(
  #Track sector simulations
  ## Track sector data
  'sector_distance' = 1000,
  'sector_velocity_kmh' = runif(n_runs,kph_min,kph_max),
  'sector_velocity_ms' = sector_velocity_kmh*1000/3600,
  
  # Is it not better to use sector_velocity as the starting point always? and the sector end velocity always as the end velocity which can result in acc or decc in the sector for x distance of this meneuver?
  
  'sector_start_velocity_kmh' = runif(n_runs,kph_min,kph_max),
  'sector_start_velocity_ms' = sector_start_velocity_kmh*1000/3600,
  'sector_end_velocity_kmh' = runif(n_runs,kph_min,kph_max),
  'sector_end_velocity_ms' = sector_end_velocity_kmh*1000/3600,
  'sector_underground' = "dry asphalt",
  'sector_alpha_slope' = alpha_slope,
  # 'sector_latitude' = 0 ,
  # 'sector_corner_radius' = 10,
  'sector_bank_slope' = alpha_bank_slope,
  # 'sector_corner_angle' = 180
)

## Vehicle and tyre performances per track underground
input_data <- input_data |> 
  mutate(
    ### brake deceleration constant on dry asphalt
    'c_full_brake_dry_asphalt' = 
      f_c_full_brake(grip_index_tyre = grip_index_tyre, 
                     x_correct_road = x_correct_road_dry_asphalt, 
                     c_full_brake_ref_tyre = c_full_brake_ref_tyre),
    ### brake deceleration constant on wet asphalt
    'c_full_brake_wet_asphalt' = 
      f_c_full_brake(grip_index_tyre = grip_index_tyre, 
                     x_correct_road = x_correct_road_wet_asphalt, 
                     c_full_brake_ref_tyre = c_full_brake_ref_tyre),
    ### linear increase before reaching optimal slip on dry asphalt
    'x_slip_long_force_dry_asphalt' = 
      f_x_slip_long_force(mu_max = 
                            ###peak friction coefficient on dry asphalt
                            f_mu_max(grip_index_tyre = grip_index_tyre,
                                     x_correct_road = x_correct_road_dry_asphalt, 
                                     mu_max_ref_tyre = mu_max_ref_tyre), 
                          ### Normal load force f 
                          normal_load_force = 
                            f_normal_load_force(m_vehicle = m_vehicle, 
                                                grav_constant = grav_constant), 
                          optimal_slip = optimal_slip_dry_asphalt),
    ### linear increase before reaching optimal slip on wet asphalt
    'x_slip_long_force_wet_asphalt' = 
      f_x_slip_long_force(mu_max =  
                            ###peak friction coefficient on wet asphalt  
                            f_mu_max(grip_index_tyre = grip_index_tyre,
                                     x_correct_road = x_correct_road_wet_asphalt, 
                                     mu_max_ref_tyre = mu_max_ref_tyre), 
                          ### Normal load force f 
                          normal_load_force = 
                            f_normal_load_force(m_vehicle = m_vehicle, 
                                                grav_constant = grav_constant), 
                          optimal_slip = optimal_slip_wet_asphalt)
  )

```

### calculate output

```{r calculate output}



ProbY <- input_data |>
  mutate(
    'firc_work_acc_dry' = f_acc_dry(sector_velocity_ms=sector_velocity_ms,
                                    sector_start_velocity_ms=sector_start_velocity_ms,
                                    frac_driver_comfort_max_accelaration=frac_driver_comfort_max_accelaration,
                                    c_accel_max=c_accel_max,
                                    m_vehicle=m_vehicle,
                                    m_rotate=m_rotate,
                                    c_drag=c_drag,
                                    A_vehicle=A_vehicle,
                                    rho_air=rho_air,
                                    v_wind=v_wind,
                                    c_roll=c_roll,
                                    grav_constant=grav_constant,
                                    sector_alpha_slope=sector_alpha_slope,
                                    optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                    x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt,
                                    accel_force_end_maneuver_1 = 0
    )$friction_work_longitude,
    'distance_acc_dry' = f_acc_dry(sector_velocity_ms=sector_velocity_ms,
                                   sector_start_velocity_ms=sector_start_velocity_ms,
                                   frac_driver_comfort_max_accelaration=frac_driver_comfort_max_accelaration,
                                   c_accel_max=c_accel_max,
                                   m_vehicle=m_vehicle,
                                   m_rotate=m_rotate,
                                   c_drag=c_drag,
                                   A_vehicle=A_vehicle,
                                   rho_air=rho_air,
                                   v_wind=v_wind,
                                   c_roll=c_roll,
                                   grav_constant=grav_constant,
                                   sector_alpha_slope=sector_alpha_slope,
                                   optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                   x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt,
                                   accel_force_end_maneuver_1 = 0
    )$distance,
    'firc_work_dec_dry' = f_dec_dry (sector_velocity_ms=sector_velocity_ms,
                                     sector_end_velocity_ms=sector_end_velocity_ms,
                                     c_decel_maneuver_2 = c_decel_comfort,
                                     m_vehicle=m_vehicle,
                                     m_rotate=m_rotate,
                                     c_drag=c_drag,
                                     A_vehicle=A_vehicle,
                                     rho_air=rho_air,
                                     v_wind=v_wind,
                                     c_roll=c_roll,
                                     grav_constant=grav_constant,
                                     c_full_brake_dry_asphalt=c_full_brake_dry_asphalt,
                                     sector_alpha_slope=sector_alpha_slope,
                                     optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                     x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt)$friction_work_longitude,
    'distance_dec_dry' = f_dec_dry (sector_velocity_ms=sector_velocity_ms,
                                    sector_end_velocity_ms=sector_end_velocity_ms,
                                    c_decel_maneuver_2 = c_decel_comfort,
                                    m_vehicle=m_vehicle,
                                    m_rotate=m_rotate,
                                    c_drag=c_drag,
                                    A_vehicle=A_vehicle,
                                    rho_air=rho_air,
                                    v_wind=v_wind,
                                    c_roll=c_roll,
                                    grav_constant=grav_constant,
                                    c_full_brake_dry_asphalt=c_full_brake_dry_asphalt,
                                    sector_alpha_slope=sector_alpha_slope,
                                    optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                    x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt)$distance,
    'firc_work_const_dry' = f_const_dry (sector_velocity_ms=sector_velocity_ms,
                                         sector_end_velocity_ms=sector_end_velocity_ms,
                                         m_vehicle=m_vehicle,
                                         m_rotate=m_rotate,
                                         c_drag=c_drag,
                                         A_vehicle=A_vehicle,
                                         rho_air=rho_air,
                                         v_wind=v_wind,
                                         c_roll=c_roll,
                                         grav_constant=grav_constant,
                                         sector_alpha_slope=sector_alpha_slope,
                                         c_full_brake_dry_asphalt=c_full_brake_dry_asphalt,
                                         optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                         x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt,
                                         distance_maneuver_3 = pmax(0,(sector_distance - 
                                                                         distance_acc_dry - 
                                                                         distance_dec_dry)))$friction_work_longitude,
    'distance_const_dry' = f_const_dry (sector_velocity_ms=sector_velocity_ms,
                                        sector_end_velocity_ms=sector_end_velocity_ms,
                                        m_vehicle=m_vehicle,
                                        m_rotate=m_rotate,
                                        c_drag=c_drag,
                                        A_vehicle=A_vehicle,
                                        rho_air=rho_air,
                                        v_wind=v_wind,
                                        c_roll=c_roll,
                                        grav_constant=grav_constant,
                                        sector_alpha_slope=sector_alpha_slope,
                                        c_full_brake_dry_asphalt=c_full_brake_dry_asphalt,
                                        optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                        x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt,
                                        distance_maneuver_3 = pmax(0,(sector_distance - 
                                                                        distance_acc_dry - 
                                                                        distance_dec_dry)))$distance,
    'firc_work_long_corn_dry' = f_corn_dry (sector_velocity_ms=sector_velocity_ms,
                                            sector_corner_radius=r_corner,
                                            m_vehicle=m_vehicle,
                                            m_rotate=m_rotate,
                                            sector_bank_slope=sector_bank_slope,
                                            sector_corner_angle=angle_corner,
                                            c_drag=c_drag,
                                            A_vehicle=A_vehicle,
                                            rho_air=rho_air,
                                            v_wind=v_wind,
                                            c_roll=c_roll,
                                            sector_alpha_slope=sector_alpha_slope,
                                            grav_constant=grav_constant,
                                            c_full_brake_dry_asphalt=c_full_brake_dry_asphalt,
                                            optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                            x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt)$friction_work_longitude,
    'firc_work_lat_corn_dry' = f_corn_dry (sector_velocity_ms=sector_velocity_ms,
                                           sector_corner_radius=r_corner,
                                           m_vehicle=m_vehicle,
                                           m_rotate=m_rotate,
                                           sector_bank_slope=sector_bank_slope,
                                           sector_corner_angle=angle_corner,
                                           c_drag=c_drag,
                                           A_vehicle=A_vehicle,
                                           rho_air=rho_air,
                                           v_wind=v_wind,
                                           c_roll=c_roll,
                                           sector_alpha_slope=sector_alpha_slope,
                                           grav_constant=grav_constant,
                                           c_full_brake_dry_asphalt=c_full_brake_dry_asphalt,
                                           optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                           x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt)$friction_work_latitude,
    'distance_corn_dry' = f_corn_dry (sector_velocity_ms=sector_velocity_ms,
                                      sector_corner_radius=r_corner,
                                      m_vehicle=m_vehicle,
                                      m_rotate=m_rotate,
                                      sector_bank_slope=sector_bank_slope,
                                      sector_corner_angle=angle_corner,
                                      c_drag=c_drag,
                                      A_vehicle=A_vehicle,
                                      rho_air=rho_air,
                                      v_wind=v_wind,
                                      c_roll=c_roll,
                                      sector_alpha_slope=sector_alpha_slope,
                                      grav_constant=grav_constant,
                                      c_full_brake_dry_asphalt=c_full_brake_dry_asphalt,
                                      optimal_slip_dry_asphalt=optimal_slip_dry_asphalt,
                                      x_slip_long_force_dry_asphalt=x_slip_long_force_dry_asphalt)$distance,
    
    .keep = "none"
  )

ProbY |> summary()

```

### prepare output data for sensitivity analysis

#### GSA of all maneuvers (total and seperate)

Sum up all forces and distances- for the 4 maneuvers and calculate output as total friction force per distance This is a scenario where the track is 10km long and there is 1 corner and 1 accelation or desceleration event.

```{r data prep}

ProbY <- ProbY |> 
  mutate(
    'Total FrWrk factor (J/m)' = (firc_work_acc_dry + 
                                    firc_work_dec_dry +
                                    firc_work_const_dry   +
                                    firc_work_long_corn_dry +
                                    firc_work_lat_corn_dry) / (distance_acc_dry              +
                                                                 distance_dec_dry              +
                                                                 distance_const_dry            +
                                                                 distance_corn_dry             ),
    'total distance (m)' = (distance_acc_dry              +
                              distance_dec_dry              +
                              distance_const_dry            +
                              distance_corn_dry             ),
    'Acceleration FrWrk factor (J/m)' = firc_work_acc_dry/distance_acc_dry,
    'Deceleration FrWrk factor (J/m)' = firc_work_dec_dry/distance_dec_dry,
    'Constant FrWrk factor (J/m)' = firc_work_const_dry/distance_const_dry,
    'Cornering FrWrk factor (J/m)' = (firc_work_long_corn_dry+firc_work_lat_corn_dry)/distance_corn_dry,
    'Cornering FrWrk (J)' = (firc_work_long_corn_dry+firc_work_lat_corn_dry)
  )

ProbY <- ProbY |> 
  rename(
    `Acceleration FrWrk (J)` = firc_work_acc_dry,
    `Deceleration FrWrk (J)` = firc_work_dec_dry,
    `Constant FrWrk (J)` = firc_work_const_dry
  )


probX <- input_data |> select(
  # remova collumns that merely intermediates from other varied variables, thereby cluttering the gsa factors
  !(c(alpha_slope, alpha_bank_slope, c_full_brake_ref_tyre, c_accel_max, p_min_acc, p_max_acc,
      sector_velocity_ms, sector_start_velocity_ms, sector_end_velocity_ms, sector_bank_slope,
      sector_underground, sector_alpha_slope,x_slip_long_force_wet_asphalt, c_full_brake_wet_asphalt  ))
) |> select(
  # select only input's that are actually varied, otherwise gsa will throw errors.
  where(~ var(.x) != 0)
)

# function to simplify the GSA for each manuevre

GSA_helper <- function(probX, ProbY, FrWrk_type){
  
  
  
  probXY <- bind_cols(
    ProbY |> select(all_of(FrWrk_type)) |> mutate(across(all_of(FrWrk_type), ~ na_if(.x, 0)))
    # transform inputs by adding a constant and taking the logarithm
    ,  probX |> mutate(across(everything(), ~log(.x+100))))  |> 
    # transform output taking logarithm
    mutate( across(all_of(FrWrk_type),~  log(.x))) |> 

    drop_na()
  
  PROB_X= select(probXY,-any_of(FrWrk_type))
  PROB_Y=  pull(probXY,all_of(FrWrk_type))
  
  m1 <- sensiFdiv(model = NULL, X=PROB_X, fdiv = "TV", nboot = 0, conf = 0.95, scale = TRUE)
  tell(m1, y=PROB_Y, S)
  
  borg_df <- tibble(colnames(PROB_X),m1$S$original)
  names(borg_df)<-c("Factor", "delta")
  borg_df$Scenario<- FrWrk_type
  
  borg_df
  
}

probX |> summary()

```

### Results



```{r data total}
# registerDoParallel(10)  # use multicore, set to the number of ou

FrWrk_type_combi <- c(c('Total FrWrk factor (J/m)','Acceleration FrWrk factor (J/m)','Deceleration FrWrk factor (J/m)', 'Constant FrWrk factor (J/m)','Cornering FrWrk factor (J/m)','Cornering FrWrk (J)','Acceleration FrWrk (J)','Deceleration FrWrk (J)','Constant FrWrk (J)'))

GSA_output1 <- GSA_helper (probX=probX, ProbY=ProbY, 
                           FrWrk_type=FrWrk_type_combi[1])
for (i in 2:9){
  GSA_output1 <- bind_rows(GSA_output1,
                           GSA_helper (probX=probX, ProbY=ProbY, 
                                       FrWrk_type=FrWrk_type_combi[i]))
}


png(filename = paste(getwd(),"/figures/Figure sensitivity analysis v2.png",sep=""),
    width = 10*0.9, height = 10*0.9, units = "in", res = 300, #pointsize = 50,
    bg = "white")

ggplot(GSA_output1, aes(x=Scenario, y=Factor, fill= delta)) + 
  geom_tile()+labs(x = "Type of friction work")+
  theme(axis.text.y = element_text(size = 8),axis.text.x = element_text(size = 8,angle = 45,hjust = 1)) 


dev.off()

# borg_df <- tibble('Factor'=NULL,"delta"=NULL)

# GSA_output <- foreach(FrWrk_type_d=FrWrk_type_combi,
#                       .combine = bind_rows,
#                       .init = ) %dopar%  {
#                         
#                         GSA_helper (probX=probX, ProbY=ProbY, 
#                                     FrWrk_type=FrWrk_type_d)
#                       }
# 
# stopImplicitCluster()
ggplot(GSA_output1, aes(x=Scenario, y=Factor, fill= delta)) + 
  geom_tile()+labs(x = "Type of friction work")+
  theme(axis.text.y = element_text(size = 8),axis.text.x = element_text(size = 8,angle = 45,hjust = 1)) 

```
