---
title: "Estimating the abrasion coefficient"
author: "Joris T.K. Quik & J. Meesters"
output: md_document
editor_options: 
  chunk_output_type: console
---

# Estimating the tyre abrasion coefficient

1. First data on the maneuvers, track, vehicle, tyres and abrassion measurements need to be combined into a dataset for use further calculations.
Then the following calculations are performed:
2. Total Force at all the tyres together
3. Total Slip at all tyres together
4. Calculate total Friction Work for the relevant abrassion measurement
5. Calculate the Abrasion Coefficient

## 1. Data prepartion
```{r data_Force_Slip_calculations}
library(tidyverse)
source("R/Base functions.R")
local_path <- "data/TWP emission data_IDIADA_v01_andSIML.xlsx"

n.Runs = 1000 # for clearly uncertrain variables, amount of values to use

source("R/SIML_dataprep.R")
source("R/SIML_force.R")
source("R/SIML_slip.R")

# AllData[1921,"Underground"]

### friction work ###
AllData_fw <- 
  AllData |> mutate(
    DistanceManAccel =
      f_accel_distance(v_start = `Start speed (m/s)`,
                       v_end = `End speed (m/s)`,
                       c_accel = `Acceleration constant (m.s^-2)`),
    DistanceManDecel = 
      f_decel_distance(v_start = `Start speed (m/s)`,
                       v_end = `End speed (m/s)`,
                       c_decel = `Deceleration constant (m.s^-2)`),
    DistanceCorn = f_corner_distance(r_corner = `Corner radius (m)`,
                                     corner_angle = `Corner angle (degrees)`),
  )  |> mutate(DistanceConst = `Distance (m)`-
                 DistanceManAccel*`Maneuver repeats` - 
                 DistanceManDecel*`Maneuver repeats`,
               FWAccelLong = list(ForceAccelLong*SlipAccelLong*DistanceManAccel*`Maneuver repeats`),
               FWDecelLong = list(ForceDecelLong*SlipDecelLong*DistanceManDecel*`Maneuver repeats`),
               FWConstLong = list(ForceConstLong*SlipConstLong*DistanceConst),
               FWLat = list(
                 SlipCornLatt_M*ForceCornLatt_M*(DistanceManAccel+DistanceManDecel) +
                   SlipCornLatt_C*ForceCornLatt_C*DistanceConst),
               FricWork_p_sector = list((FWAccelLong+FWDecelLong+FWConstLong+FWLat)) # calculate FW per sector
               
  )



```


## 4. Total Friction Work

```{r FrictionWork totals}

FWtotals <-
  AllData_fw |> ungroup() |> 
  group_by(AbrasionTest,Tyre_brand,`Vehicle type`,Scenario,`Sector number`,`Maneuver number`,`Test section`) |> 
  unnest(FricWork_p_sector) #get the uncertainty runs to the tibble for summarise


length(FWtotals$Track)
FWtotals <- FWtotals |>  mutate(RUN = rep(1:n.Runs)) # give code to each run

FWtotals <-
  FWtotals |> ungroup() |> 
  group_by(AbrasionTest,Tyre_brand,`Vehicle type`,Scenario,`Test section`,RUN,`Total distance (km)`, `Number of laps`) |> 
  summarise(sector_count = n(),
            PartFrictionWork = sum(FricWork_p_sector)) |> # sum up all friction work per test section(for maneuvers?) (e.g. rural 60 kph)
  mutate(SectionFrictionWork = PartFrictionWork *`Number of laps`) # multiple test section with number of laps to get total FW per test per section

# write.csv(FWtotals, "FWtotals2.csv")

FWtotals <-
  FWtotals |>  ungroup() |> 
  group_by(AbrasionTest,Tyre_brand,`Vehicle type`,Scenario,RUN) |> 
  summarise(Section_count = n(),
            TotFrictionWork = sum(SectionFrictionWork), # sum the test sections to get total FW for the whole arbasion test
            Test_Distance_km = sum(`Total distance (km)`)) |> # sum up also the distance of each test section to get total distace driven in each test.
  mutate(FW_p_km = TotFrictionWork/Test_Distance_km) # devide total FW per test by the distance driven per test

library(ggplot2)

plotdata <- FWtotals |> filter(Tyre_brand == "Dunlop" & Scenario == "Standard") # plot FW for each test (J per km)

plot <- ggplot(plotdata, aes(x = c(AbrasionTest), y = FW_p_km, fill=Tyre_brand )) +
  geom_violin()
plot+  scale_y_log10(labels = scales::number_format())
# boxplot(FWtotals)


```

```{r}
plotdata <- FWtotals |> filter( Scenario == "Standard") # plot FW for each test (J per km)

plot <- ggplot(plotdata, aes(x = c(AbrasionTest), y = FW_p_km, fill=Tyre_brand )) +
  geom_violin()
plot+  scale_y_log10(labels = scales::number_format())


plotdata <- FWtotals |> filter( Scenario == "High load"|Scenario == "Standard",Tyre_brand == "LingLong") # plot FW for each test (J per km)

plot <- ggplot(plotdata, aes(x = c(AbrasionTest), y = FW_p_km, fill=Scenario )) +
  geom_violin()
plot+  scale_y_log10(labels = scales::number_format())

### end of FW

```

## 5. Abrasion Coefficient

```{r abrasion coeff}

# read in and prepare abrasion rate measurement data
IDIADAwear <- readxl::read_excel(path = local_path, sheet = "Abrasion", skip = 9)

#provide proper column names
names(IDIADAwear)[c(1,2)] <- c("Tyre_brand","Wheel")

# extend Tyre to all its rows
for(i in 1:nrow(IDIADAwear)){
  if(is.na(IDIADAwear$Tyre_brand[i])){
    IDIADAwear$Tyre_brand[i] <- IDIADAwear$Tyre_brand[i-1]
  }
}

#What?
IDIADAwear$C0 <- NULL

# clean up rows with NA and calculated data
IDIADAwear <- 
  IDIADAwear |> separate(col = Tyre_brand,
                         into =c("Tyre_brand", "Scenario"),
                         sep = "     ")|> 
  filter(!is.na(Wheel)) |> 
  filter(Wheel != "Total") |> 
  filter( Wheel != "Front") |> 
  filter(Wheel != "Rear") 

#Track to long format
WearAsLong <- tidyr::pivot_longer(IDIADAwear, 
                                  cols = names(IDIADAwear)[!names(IDIADAwear) %in% c("Tyre_brand","Wheel","Scenario")],
                                  names_to = c("Shift","AbrasionTest"),
                                  names_sep = "-",
                                  values_to = "Abrasion_mg_km")

WearAsLong <-
  WearAsLong |> mutate(
    AbrasionTest = 
      AbrasionTest |> recode(
        "Runin" = "Run-in",
        "Rur" = "Rural",
        "Mot" = "Motorway",
        "Ur" = "Urban"
      ),
    Scenario = Scenario |> 
      replace_na("Standard")
  )

WearAsLong <- # aggregate data per wheel to total per vehicle
  WearAsLong |> ungroup() |> 
  group_by(Tyre_brand ,Scenario ,Shift , AbrasionTest) |> 
  summarise(Abrasion_mg_km = sum(Abrasion_mg_km),
            nWheels = n())

CombinedFW_TW <- 
  left_join(WearAsLong,FWtotals, relationship = "many-to-many") |> 
  mutate(AbrasionCoeff = Abrasion_mg_km / FW_p_km) |> # Calculate the abrasion coefficient
  mutate(TestScenName = paste(Tyre_brand, Scenario, sep = "-"))

ABR_COEF_avg <-
  CombinedFW_TW |> filter(AbrasionTest != "Run-in" & Scenario == "Standard") |> ungroup() |> group_by(AbrasionTest) |>  summarise(mean(AbrasionCoeff))

unique(CombinedFW_TW$AbrasionTest)
# figure of tyre fricction work

kData25 <-  CombinedFW_TW |> filter(AbrasionTest != "Run-in" & Scenario == "Standard") 

# str(kData25)

```


```{r figures standard scenarios}
plot_theme = theme(
  axis.title.x = element_text(size = 16),
  axis.text = element_text(size = 14), 
  axis.title.y = element_text(size = 16),
  plot.background = element_rect(fill = 'white'),
  panel.background = element_rect(fill = 'white'),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color='black'),
  plot.margin = margin(1, 1, 1, 1, "cm")
  #panel.grid.major = element_line(colour = "grey",size=0.25)
)

kData25

FW_plot <- ggplot(kData25, 
                  aes(x = reorder(AbrasionTest,FW_p_km), y = FW_p_km/1000, fill=reorder(AbrasionTest, FW_p_km))) +
  geom_violin()+
  theme(legend.position="none")+
  # scale_y_log10(labels = scales::number_format()) +             # Log-transform y-axis
  labs(x = "Abrasion test scenario", y = "Friction Work (J/m)") +                   # Adjust labels
  # coord_flip() +
  plot_theme +
  labs(fill = "Tyre")
FW_plot

ggsave(paste0("figures/FrictionWork_AbrasionT_Tyre_5x5",format(Sys.time(),'%Y%m%d'),".png"),
       width = 5, height = 5)


# Figure of abrasion measurements per tyre type tested
plotdata <- CombinedFW_TW |> 
  filter( Scenario == "Standard") |> ungroup() |> 
  group_by(Tyre_brand ,Scenario , Shift, AbrasionTest , `Vehicle type`,TestScenName) |>   summarise(Abrasion_mg_km=mean(Abrasion_mg_km),
                                                                                                    n=n(),
                                                                                                    sd=sd(Abrasion_mg_km))

plotdata |> ungroup() |> 
  group_by(Scenario , AbrasionTest , `Vehicle type`) |> 
  summarise(            min = min(Abrasion_mg_km),
                        p5 = quantile(Abrasion_mg_km, probs=0.05),
                        AvgAbrasion_mg_km = mean(Abrasion_mg_km),
                        p50 = quantile(Abrasion_mg_km, probs=0.5),
                        p95 = quantile(Abrasion_mg_km, probs=0.95),
                        max = max(Abrasion_mg_km),
                        
                        n_abr = n())
ggplot(kData25, 
       aes(x = reorder(AbrasionTest, Abrasion_mg_km), y = Abrasion_mg_km, fill = AbrasionTest ) ) +
  geom_violin() + theme(legend.position="none")+
  # geom_point(aes(colour = reorder(TestScenName, Abrasion_mg_km)), 
  # position = position_jitterdodge(jitter.width = 0.1,
  # dodge.width = 0.75))+
  labs(x = "Abrasion test scenario", y = "Abrasion rate (mg/km)") +                   # Adjust labels
  
  plot_theme 

ggsave(paste0("figures/AbrasionRate_AbrasionT_Tyre_5x5",format(Sys.time(),'%Y%m%d'),".png"),
       width = 5, height = 5)

kData25_2 <- kData25 |> 
  mutate(AbrasionTest = factor(AbrasionTest, levels = c("Rural", "Motorway", "Urban")))

ACplot1 <- ggplot(kData25_2, 
                  aes(x = c(AbrasionTest), y = AbrasionCoeff*1000,fill = AbrasionTest )) +
  geom_violin() +theme(legend.position="none")+
  # geom_point(aes(colour = reorder(TestScenName, AbrasionCoeff)), 
  # position = position_jitterdodge(jitter.width = 0.05,
  # dodge.width = 0.75))+
  labs(x = "Abrasion test scenario", y = "Abrasion Coeff (mg/kJ)") +                   # Adjust labels
  
  plot_theme 
ACplot1

ggsave(paste0("figures/AbrasionCoef1_AbrasionT_Tyre_5x5",format(Sys.time(),'%Y%m%d'),".png"),
       width = 5, height = 5)


# Normalized_motorway <- 
FWtotals |>ungroup() |>  distinct(AbrasionTest)
Normalized_motorway <-
  FWtotals |>ungroup() |>  filter(AbrasionTest == "Motorway_SIML"&Scenario == "Standard") |> rename(Normalized_FW_p_km = FW_p_km) |> select(Tyre_brand, `Vehicle type`,Scenario,Normalized_FW_p_km) |> mutate(RUN = rep.int(rep(1:n.Runs),5))

Normalized_motorway2 <-
  kData25 |> filter(AbrasionTest == "Motorway") |> 
  left_join(Normalized_motorway) |> 
  mutate(MotorWayEF_mgkm = Normalized_FW_p_km*AbrasionCoeff) |> 
  mutate(AbrasionTest = "Motorway sim")

Normalized_motorway2 |> ungroup() |> distinct(`Vehicle type`)

ACplot2 <- ggplot(Normalized_motorway2, 
                  aes(x = reorder(Tyre_brand,MotorWayEF_mgkm), y = MotorWayEF_mgkm ,fill = Tyre_brand)) +
  geom_violin() +theme(legend.position="none") +
  # geom_point(aes(colour = TestScenName), position = position_jitterdodge(jitter.width = 0.05,
  #                                                                      dodge.width = 0.75))+
  labs(x = "Abrasion test scenario", y = "Abrasion Rate (mg/km)") +                   # Adjust labels
  
  plot_theme 
ACplot2
ggsave(paste0("figures/EmissionFactor_sim_Tyre",format(Sys.time(),'%Y%m%d'),".png"),
       width = 5, height = 5)

```


### ADAC
Add Abrasion data from ADAC

```{r ADAC}

ADAC_wear <- readxl::read_excel(path = local_path, sheet = "Abrasion_ADAC")

Normalized_motorway3 <-
  ADAC_wear |> mutate(AbrasionTest = "ADAC") |> 
  rename(
    MotorWayEF_mgkm = `Abrasion rate (mg/km)`,
    Tyre_brand = Tyre
  ) |> full_join(Normalized_motorway2)


plot <- ggplot(Normalized_motorway3, 
               aes(x = reorder(AbrasionTest,MotorWayEF_mgkm), y = MotorWayEF_mgkm ,fill = AbrasionTest)) +
  geom_violin() + theme(legend.position="none") +
  # geom_point(aes(colour = TestScenName), position = position_jitterdodge(jitter.width = 0.05,
  #                                                                      dodge.width = 0.75))+
  labs(x = "Abrasion test scenario", y = "Abrasion Rate (mg/km)") +                   # Adjust labels
  
  plot_theme 
plot
ggsave(paste0("figures/EmissionFactor_sim_ADAC",format(Sys.time(),'%Y%m%d'),".png"),
       width = 4, height = 5)

```

### Celik data

```{r celik}

Celik_AC <- readxl::read_excel(path = local_path, sheet = "Cenek et al. 1997")
# kData25

Celik_AC <-
  Celik_AC |> rename(
    AbrasionCoeff = `abrasion coefficient (g / MJ)`
  ) |> mutate(AbrasionCoeff = AbrasionCoeff/1000,
              AbrasionTest = "Celik_1997", .keep = "used") |> 
  full_join(kData25) |> 
  mutate(AbrasionTest = factor(AbrasionTest, levels = c("Rural", "Motorway", "Urban","Celik_1997")))



ACplot1 <- ggplot(Celik_AC, 
                  aes(x = c(AbrasionTest), y = AbrasionCoeff*1000,fill = AbrasionTest )) +
  geom_violin() +theme(legend.position="none")+
  # geom_point(aes(colour = reorder(TestScenName, AbrasionCoeff)), 
  # position = position_jitterdodge(jitter.width = 0.05,
  # dodge.width = 0.75))+
  labs(x = "Abrasion test scenario", y = "Abrasion Coeff (mg/kJ)") +                   # Adjust labels
  
  plot_theme 
ACplot1

ggsave(paste0("figures/AbrasionCoef_celik_5x5",format(Sys.time(),'%Y%m%d'),".png"),
       width = 6, height = 5)



```


