---
title: "Estimating the abrasion coefficient"
author: "Joris T.K. Quik"
output: md_document
editor_options: 
  chunk_output_type: console
---

# Estimating the tyre abrasion coefficient

1. First data on the maneuvers, track, vehicle, tyres and abrassion measurements need to be combined into a dataset for use further calculations.
Then the following calculations are performed:
2. Total Force at all the tyres together
3. Total Slip at all tyres together
4. Calculate total Friction Work for the relevant abrassion measurement
5. Calculate the Abrasion Coefficient

## 1. Data prepartion
```{r data}
library(tidyverse)
source("R/Base functions.R")
local_path <- "data/TWP emission data_IDIADA_v01.xlsx"

n.Runs = 1000 # for clearly uncertrain variables, amount of values to use

Maneuver_data <- readxl::read_excel(path = local_path, sheet = "Maneuver data")

AllData <- 
  Maneuver_data |> 
  full_join(readxl::read_excel(path = local_path, sheet = "Sector data")) |> 
  cross_join(readxl::read_excel(path = local_path, sheet = "Vehicle data")) |> 
  full_join(readxl::read_excel(path = local_path, sheet = "Tyre_data")) |> 
  full_join(readxl::read_excel(path = local_path, sheet = "Scenario data"))

Constants <- readxl::read_excel(path = local_path, sheet = "Constants")
# str(AllData)
# AllData$Vehicle_class

Tyre_Label_table_fuelEff =    readxl::read_excel(
  "data/Tyre_conversion.xlsx", 
  sheet = "Label fuel efficiency class",
  skip=1)
Tyre_Label_table_wetgrip =    readxl::read_excel(
  "data/Tyre_conversion.xlsx", 
  sheet = "Label wet grip class",
  skip=1)

AllData2 <-
  AllData |> 
  rowwise() |>  
  mutate(
    RolCoef_min =  fRolCoef_Tlabel(
      Label_fuelleff =
        `Fuel efficiency class`,
      Vehicle_class = Vehicle_class,
      Tyre_Label_table =
        Tyre_Label_table_fuelEff
    )$min,
    RolCoef_max = fRolCoef_Tlabel(
      Label_fuelleff =
        `Fuel efficiency class`,
      Vehicle_class = Vehicle_class,
      Tyre_Label_table = Tyre_Label_table_fuelEff
    )$max,
    GripIndex_min  = fGripIndex_Tlabel(
      Label_wetgrip = `Wet grip class`,
      Vehicle_class = Vehicle_class,
      Tyre_Label_table = Tyre_Label_table_wetgrip
    )$min,
    GripIndex_max = fGripIndex_Tlabel(
      Label_wetgrip = `Wet grip class`,
      Vehicle_class = Vehicle_class,
      Tyre_Label_table = Tyre_Label_table_wetgrip
    )$max
  )
# create n.Runs number of Rol Coefficients based on uncertainty of tyre label rol coefficient classes
AllData3 <-
  AllData2 |> mutate(
    RolCoef_u = list(runif(n.Runs, RolCoef_min, RolCoef_max)),
    GripIndex_u = list(runif(n.Runs, GripIndex_min, GripIndex_max)),
    
    x_correct_mu_max_track_u =
      switch(
        Underground,
        "Wet asphalt" = 1,
        "Dry asphalt" = list(
          runif(
            n.Runs,
            min =  Constants |>
              filter(Name == "x_correct_mu_max_wet2dry_min") |>
              pull(value),
            max =  Constants |>
              filter(Name == "x_correct_mu_max_wet2dry_max") |>
              pull(value)
          )
        )
      ),
    
    optimal_slip_ratio_track_u =
      switch(
        Underground,
        "Wet asphalt" = list(
          runif(
            n.Runs,
            min =  Constants |>
              filter(Name == "optimal_slip_wet_min") |>
              pull(value),
            max =  Constants |>
              filter(Name == "optimal_slip_wet_max") |>
              pull(value)
          )
        ),
        "Dry asphalt" = list(
          runif(
            n.Runs,
            min =  Constants |>
              filter(Name == "optimal_slip_dry_min") |>
              pull(value),
            max =  Constants |>
              filter(Name == "optimal_slip_dry_max") |>
              pull(value)
          )
        )
      )
  )


### end data prep ###


```


## 2. Total Force
Longitutidal and Lattidunal forces are calculated

```{r force}
Force <-
  AllData3 |> mutate(
    ForceDecelLong =
      list(
        f_decel_long_force(
          c_roll = RolCoef_u,
          m_vehicle = `Mass (kg)`,
          grav_constant = Constants |>
            filter(Name == "grav_constant") |> pull(value),
          c_drag = `Aero_drag_coef (-)`,
          A_vehicle = `Surface_Area (m2)`,
          rho_air = Constants |>
            filter(Name == "rho_air") |> pull(value),
          v_start_decel = `Start speed (m/s)`,
          v_end_decel = `End speed (m/s)`,
          v_wind = v_wind,
          alpha_slope = `Longitudinal slope (%)` / 100,
          m_rotate = runif(
            n.Runs,
            Constants |>
              filter(Name == "min_rotating_fraction") |> pull(value),
            Constants |>
              filter(Name == "max_rotating_fraction") |> pull(value)
          ),
          c_decel = `Deceleration constant (m.s^-2)`
        )
      ),
    ForceAccelLong =
      list(
        f_accel_long_force(
          c_roll = RolCoef_u,
          m_vehicle = `Mass (kg)`,
          grav_constant = Constants |>
            filter(Name == "grav_constant") |> pull(value),
          c_drag = `Aero_drag_coef (-)`,
          A_vehicle = `Surface_Area (m2)`,
          rho_air = Constants |>
            filter(Name == "rho_air") |> pull(value),
          v_start_accel = `Start speed (m/s)`,
          v_end_accel = `End speed (m/s)`,
          v_wind = v_wind,
          alpha_slope = `Longitudinal slope (%)` / 100,
          m_rotate = runif(
            n.Runs,
            Constants |>
              filter(Name == "min_rotating_fraction") |> pull(value),
            Constants |>
              filter(Name == "max_rotating_fraction") |> pull(value)
          ),
          c_accel = `Acceleration constant (m.s^-2)`
        )
      ),
    ForceConstLong =
      list(
        f_const_speed_long_force(
          c_roll = RolCoef_u,
          m_vehicle = `Mass (kg)`,
          grav_constant = Constants |>
            filter(Name == "grav_constant") |> pull(value),
          c_drag = `Aero_drag_coef (-)`,
          A_vehicle = `Surface_Area (m2)`,
          rho_air = Constants |>
            filter(Name == "rho_air") |> pull(value),
          v_vehicle = `End speed (m/s)`,
          v_wind = `v_wind`,
          alpha_slope = `Longitudinal slope (%)` /100)
      ),
    ForceCornLatt_M =
      list(
        f_lat_force( m_vehicle = `Mass (kg)`,
                     grav_constant = Constants |> 
                       filter(Name == "grav_constant") |> pull(value),
                     r_corner = `Corner radius (m)`,
                     alpha_bank_slope = `Latitudinal slope (%)` / 100,
                     v_vehicle = mean(`Start speed (m/s)`,`End speed (m/s)`)) 
      ),
    ForceCornLatt_C =
      list(
        f_lat_force( m_vehicle = `Mass (kg)`,
                     grav_constant = Constants |> 
                       filter(Name == "grav_constant") |> pull(value),
                     r_corner = `Corner radius (m)`,
                     alpha_bank_slope = `Latitudinal slope (%)` / 100,
                     v_vehicle = `End speed (m/s)`) 
      )
  )

```

## 3. Total Slip

```{r slip}
# add slipt to the data:

Slip <-
  Force |> mutate(
    SlipDecelLong =
      list(
        f_decel_long_slip(c_roll = RolCoef_u, 
                          m_vehicle = `Mass (kg)`, 
                          grav_constant = Constants |> 
                            filter(Name == "grav_constant") |> pull(value), 
                          c_drag = `Aero_drag_coef (-)`, 
                          A_vehicle = `Surface_Area (m2)`,
                          rho_air = Constants |> 
                            filter(Name == "rho_air") |> pull(value),
                          v_start_decel = `Start speed (m/s)`, 
                          v_end_decel = `End speed (m/s)`, 
                          v_wind = `v_wind`,
                          alpha_slope = `Longitudinal slope (%)`/100, 
                          m_rotate = runif(n.Runs,
                                           Constants |> 
                                             filter(Name == "min_rotating_fraction") |> 
                                             pull(value),
                                           Constants |> 
                                             filter(Name == "max_rotating_fraction") |> 
                                             pull(value)), 
                          c_decel = `Deceleration constant (m.s^-2)`, 
                          optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                          grip_index_tyre = GripIndex_u,
                          wet_mu_max_ref_tyre = Constants |>
                            filter(Name == "mu_max_ref_tyre_wet") |> pull(value),
                          x_correct_mu_max_track = x_correct_mu_max_track_u,
                          c_brake_ref_tyre_wet = Constants |> 
                            filter(Name == "c_brake_ref_tyre_wet") |> pull(value)
        ) 
      ),
    SlipAccelLong= list(
      f_accel_long_slip(c_roll = RolCoef_u, 
                        m_vehicle = `Mass (kg)`, 
                        grav_constant = Constants |> 
                          filter(Name == "grav_constant") |> pull(value), 
                        c_drag = `Aero_drag_coef (-)`, 
                        A_vehicle = `Surface_Area (m2)`, 
                        rho_air = Constants |> 
                          filter(Name == "rho_air") |> pull(value),
                        v_start_accel = `Start speed (m/s)`, 
                        v_end_accel = `End speed (m/s)`, 
                        v_wind = v_wind, 
                        alpha_slope = `Longitudinal slope (%)`/100, 
                        m_rotate = runif(n.Runs,
                                         Constants |> 
                                           filter(Name == "min_rotating_fraction") |> 
                                           pull(value),
                                         Constants |> 
                                           filter(Name == "max_rotating_fraction") |> 
                                           pull(value)), 
                        c_accel = `Acceleration constant (m.s^-2)`, 
                        optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                        grip_index_tyre = GripIndex_u, 
                        wet_mu_max_ref_tyre = Constants |> 
                          filter(Name == "mu_max_ref_tyre_wet") |> pull(value),
                        x_correct_mu_max_track = x_correct_mu_max_track_u)
    ),
    SlipConstLong = 
      list(
        f_const_speed_long_slip(c_roll = RolCoef_u, 
                                m_vehicle = `Mass (kg)`, 
                                m_rotate = runif(n.Runs,
                                                 Constants |> 
                                                   filter(Name == "min_rotating_fraction") |> 
                                                   pull(value),
                                                 Constants |> 
                                                   filter(Name == "max_rotating_fraction") |> 
                                                   pull(value)),
                                grav_constant = Constants |> 
                                  filter(Name == "grav_constant") |> pull(value), 
                                c_drag = `Aero_drag_coef (-)`, 
                                A_vehicle = `Surface_Area (m2)`, 
                                rho_air = Constants |> 
                                  filter(Name == "rho_air") |> pull(value), 
                                v_vehicle = `End speed (m/s)`, 
                                v_wind = `v_wind`,
                                alpha_slope = `Longitudinal slope (%)`/100,
                                wet_mu_max_ref_tyre = Constants |> 
                                  filter(Name == "mu_max_ref_tyre_wet") |> pull(value), 
                                optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                                grip_index_tyre = GripIndex_u, 
                                c_brake_ref_tyre_wet = Constants |> 
                                  filter(Name == "c_brake_ref_tyre_wet") |> pull(value), 
                                x_correct_mu_max_track = x_correct_mu_max_track_u)
      ),
    SlipCornLatt_M =
      list(
        f_lat_slip(m_vehicle = `Mass (kg)`, 
                   v_vehicle = mean(`Start speed (m/s)`,
                                    `End speed (m/s)`), 
                   r_corner = `Corner radius (m)`, 
                   grav_constant = Constants |> 
                     filter(Name == "grav_constant") |> pull(value),
                   alpha_bank_slope = `Latitudinal slope (%)`/100, 
                   optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                   grip_index_tyre = GripIndex_u, 
                   wet_mu_max_ref_tyre = Constants |> 
                     filter(Name == "mu_max_ref_tyre_wet") |> pull(value), 
                   x_correct_mu_max_track = x_correct_mu_max_track_u)
      ),
    SlipCornLatt_C =
      list(
        f_lat_slip(m_vehicle = `Mass (kg)`, 
                   v_vehicle = `End speed (m/s)`, 
                   r_corner = `Corner radius (m)`, 
                   grav_constant = Constants |> 
                     filter(Name == "grav_constant") |> pull(value),
                   alpha_bank_slope = `Latitudinal slope (%)`/100, 
                   optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                   grip_index_tyre = GripIndex_u, 
                   wet_mu_max_ref_tyre = Constants |> 
                     filter(Name == "mu_max_ref_tyre_wet") |> pull(value), 
                   x_correct_mu_max_track = x_correct_mu_max_track_u)
      ) 
  )


```

## 4. Total Friction Work

```{r FrictionWork}
TotalFW <- 
  Slip |> mutate(
    DistanceManAccel =
      f_accel_distance(v_start = `Start speed (m/s)`,
                       v_end = `End speed (m/s)`,
                       c_accel = `Acceleration constant (m.s^-2)`),
    DistanceManDecel = 
      f_decel_distance(v_start = `Start speed (m/s)`,
                       v_end = `End speed (m/s)`,
                       c_decel = `Deceleration constant (m.s^-2)`),
    DistanceCorn = f_corner_distance(r_corner = `Corner radius (m)`,
                                     corner_angle = `Corner angle (degrees)`)
  )

FrictionWork = 
  TotalFW |> mutate(DistanceConst = `Distance (m)`-
                      DistanceManAccel*`Maneuver repeats` - 
                      DistanceManDecel*`Maneuver repeats`,
                    FWAccelLong = list(ForceAccelLong*SlipAccelLong*DistanceManAccel*`Maneuver repeats`),
                    FWDecelLong = list(ForceDecelLong*SlipDecelLong*DistanceManDecel*`Maneuver repeats`),
                    FWConstLong = list(ForceConstLong*SlipConstLong*DistanceConst),
                    FWLat = list(
                      SlipCornLatt_M*ForceCornLatt_M*(DistanceManAccel+DistanceManDecel) +
                      SlipCornLatt_C*ForceCornLatt_C*DistanceConst),
                    TotalFricWork = list((FWAccelLong+FWDecelLong+FWConstLong+FWLat)*`Number of laps`),
                    .keep = c("used")
                    
  )


```


```{r abrasion data}



library(readxl)
IDIADAwear <- readxl::read_excel("data/Abrasion test_WP2.3_Leon-T_IDIADA.xlsx", skip = 9)

#provide proper column names
names(IDIADAwear)[c(1,2)] <- c("Tyre","Wheel")
names(IDIADAwear) <- sapply(names(IDIADAwear), function(x){
  gsub("-", "_", x)
})
#extend Tyre to all its rows
for(i in 1:nrow(IDIADAwear)){
  if(is.na(IDIADAwear$Tyre[i])){
    IDIADAwear$Tyre[i] <- IDIADAwear$Tyre[i-1]
  }
}
#What?
IDIADAwear$C0 <- NULL
#Track to long format
WearAsLong <- tidyr::pivot_longer(IDIADAwear, 
                           names(IDIADAwear)[!names(IDIADAwear) %in% c("Tyre","Wheel")],
                           names_to = "track",
                           values_to = "wear")
WearAsLong$track[WearAsLong$track %in% c("R1", "R2", "R3")] <- "R"

#remove the aggregates 
WearAsLong <- WearAsLong[WearAsLong$Wheel %in% c("FR ", "FL", "RR ", "RL"),]
Map2track <- unique(WearAsLong$track)
table(WearAsLong$track)




```

