---
title: "Sensitivity analysis of Tyre friction abrasion model v1.1"
author: "Joris T.K. Quik"
output: md_document
editor_options: 
  chunk_output_type: console
---

# Estimating the tyre abrasion coefficient

1. First data on the maneuvers, track, vehicle, tyres and abrassion measurements need to be combined into a dataset for use further calculations.
Then the following calculations are performed:
2. Total Force at all the tyres together
3. Total Slip at all tyres together
4. Calculate total Friction Work for the relevant abrassion measurement
5. Perform Global Sensitivity Analysis

## 1. Data prepartion
```{r data,  echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
source("R/Base functions.R")
local_path <- "data/TWP emission data_IDIADA_v01_andSIML.xlsx"

n.Runs = 1000 # for clearly uncertrain variables, amount of values to use
# set.seed = (120847)

test_data <- readxl::read_excel(path = local_path, sheet = "Test data")



Constants <- readxl::read_excel(path = local_path, sheet = "Constants")
# str(AllData)
# AllData$Vehicle_class

Vehicle_Class_override = "C1"

min_mass_vehicle = 900
max_mass_vehicle = 3000
min_A_vehicle = 2
max_A_vehicle = 4
min_c_drag_vehicle = 0.02
max_c_drag_vehicle = 0.5

min_v_wind = -5
max_v_wind = 5
min_grade_slope = -10
max_grade_slope = 10
min_bank_slope = -1
max_bank_slope = 1

min_corn_radius = 500
max_corn_radius = 1000
min_corn_angle = 45
max_corn_angle = 90


Tyre_Label_table_fuelEff =    readxl::read_excel(
  "data/Tyre_conversion.xlsx", 
  sheet = "Label fuel efficiency class",
  skip=1)
Tyre_Label_table_wetgrip =    readxl::read_excel(
  "data/Tyre_conversion.xlsx", 
  sheet = "Label wet grip class",
  skip=1)
RolCoef_min =  fRolCoef_Tlabel(
  Label_fuelleff =
    "A",
  Vehicle_class =  Vehicle_Class_override,
  Tyre_Label_table =
    Tyre_Label_table_fuelEff
)$min
RolCoef_max = fRolCoef_Tlabel(
  Label_fuelleff =
    "D",
  Vehicle_class = Vehicle_Class_override,
  Tyre_Label_table = Tyre_Label_table_fuelEff
)$max
GripIndex_min  = fGripIndex_Tlabel(
  Label_wetgrip = "D",
  Vehicle_class = Vehicle_Class_override,
  Tyre_Label_table = Tyre_Label_table_wetgrip
)$min
GripIndex_max = fGripIndex_Tlabel(
  Label_wetgrip =  "A",
  Vehicle_class =  Vehicle_Class_override,
  Tyre_Label_table = Tyre_Label_table_wetgrip
)$max


## for sensitivity only include 1 set of data for Tyre, Test and Vehicle data
AllData <-
  test_data |> 
  # distinct(`Test section`)
  filter(`Test section` == "Motorway_SIML")

AllData <-
  AllData |> 
  rowwise() |>  
  mutate(
    
    RolCoef_u = list(runif(n.Runs, RolCoef_min, RolCoef_max)),
    GripIndex_u = list(runif(n.Runs, GripIndex_min, GripIndex_max)),
    Underground = "Dry asphalt"
  )
# create n.Runs number of Rol Coefficients based on uncertainty of tyre label rol coefficient classes
AllData <-
  AllData |> mutate(
    x_correct_mu_max_track_u =
      switch(
        Underground,
        "Wet asphalt" = 1,
        "Dry asphalt" = list(
          runif(
            n.Runs,
            min =  Constants |>
              filter(Name == "x_correct_mu_max_wet2dry_min") |>
              pull(value),
            max =  Constants |>
              filter(Name == "x_correct_mu_max_wet2dry_max") |>
              pull(value)
          )
        )
      ),
    
    optimal_slip_ratio_track_u =
      switch(
        Underground,
        "Wet asphalt" = list(
          runif(
            n.Runs,
            min =  Constants |>
              filter(Name == "optimal_slip_wet_min") |>
              pull(value),
            max =  Constants |>
              filter(Name == "optimal_slip_wet_max") |>
              pull(value)
          )
        ),
        "Dry asphalt" = list(
          runif(
            n.Runs,
            min =  Constants |>
              filter(Name == "optimal_slip_dry_min") |>
              pull(value),
            max =  Constants |>
              filter(Name == "optimal_slip_dry_max") |>
              pull(value)
          )
        )
      )
  )

min_topspeed = 90 / 3.6
max_topspeed = 130 / 3.6
min_lowspeed = 0
max_lowspeed = 89 / 3.6
min_constspeed = 60 / 3.6

min_accconst = 0.1
max_accconst = 1.6677
min_decconst = 0.1
max_decconst = 2.94

AllData <- 
  AllData |> 
  mutate(
    RUNS = list(c(1:n.Runs)),
    `Mass (kg)` = list(runif(n.Runs, min_mass_vehicle, max_mass_vehicle)),
    `Aero_drag_coef (-)` = list(runif(n.Runs, min_c_drag_vehicle, max_c_drag_vehicle)),
    `Surface_Area (m2)` = list(runif(n.Runs, min_A_vehicle, max_A_vehicle)),
    `Longitudinal slope (%)` = list(runif(n.Runs,min_grade_slope,max_grade_slope)), 
    `Latitudinal slope (%)` = list(runif(n.Runs,min_bank_slope,max_bank_slope)), 
    # `Corner radius (m)` = list(runif(n.Runs,)),
    # `Corner angle (degrees)` = list(runif(n.Runs,)),
    v_wind = list(runif(n.Runs, min_v_wind, max_v_wind)),
    m_rotate = list(runif(
      n.Runs,
      Constants |>
        filter(Name == "min_rotating_fraction") |> pull(value),
      Constants |>
        filter(Name == "max_rotating_fraction") |> pull(value)
    )
    )
  )

AllData <- 
  AllData |> 
  mutate(
    v_start_decel = list(
      runif(n.Runs, min_topspeed, max_topspeed)),
    v_end_decel =  list(
      runif(n.Runs, min_lowspeed, max_lowspeed)
    ),
    v_start_accel = list(
      runif(n.Runs, min_lowspeed, max_lowspeed)
    ),
    v_end_accel =  list(
      runif(n.Runs, min_topspeed, max_topspeed)
    ),
    v_const = list(
      runif(n.Runs, min_constspeed, max_topspeed)
    ),
    `Deceleration constant (m.s^-2)` = list(
      runif(n.Runs, min_decconst, max_decconst)
    ),
    `Acceleration constant (m.s^-2)` = list(
      runif(n.Runs, min_accconst, max_accconst)
    )
  )

AllData <- 
  AllData |> mutate(
    `Corner radius (m)` =  list(runif(n.Runs,
                                      min_corn_radius,
                                      max_corn_radius)),
    `Corner angle (degrees)` = list(runif(n.Runs,
                                          min_corn_angle,
                                          max_corn_angle))
  )


PROB_X <- AllData |> 
  select(!c(`Section name`,
            `Test section`, 
            AbrasionTest,  
            Track,               
            `Number of laps`, 
            `Total distance (km)`,
            Underground))


AllData <- 
  AllData |> 
  left_join(readxl::read_excel(path = local_path, sheet = "Sector data") |> select("Track","Sector number","Sector name") ) |> 
  # cross_join(readxl::read_excel(path = local_path, sheet = "Vehicle data")) |> 
  # full_join(readxl::read_excel(path = local_path, sheet = "Tyre_data"),relationship = "many-to-many") |> 
  left_join(readxl::read_excel(path = local_path, sheet = "Maneuver data") |> select("Track","Sector number","Test section","Maneuver number", "Maneuver code","Maneuver repeats"))

AllData <-
  AllData |> mutate(
    `Sector Distance (m)` = 
      switch(`Sector name`, 
             "East BendSIML" = 
               list(2*pi*`Corner radius (m)`*`Corner angle (degrees)`/360),
             "North StraightSIML" = 
               list(`Total distance (km)`*1000 - 
                      2*pi*`Corner radius (m)`*`Corner angle (degrees)`/360)
      )
  )


```


```{r  echo=FALSE, message=FALSE, warning=FALSE}
AllData <-
  AllData |> unnest(cols = where(is.list))
AllData
```


## 2. Total Force
Longitutidal and Lattidunal forces are calculated

```{r force, echo=FALSE, message=FALSE, warning=FALSE}

AllData <-
  AllData |> 
  rowwise() |> # rowwise needed because there are ifstatements that do not work with vectors
  mutate(
    ForceLong =
      switch(
        `Maneuver code`,
        "Deceleration" = f_decel_long_force(
          c_roll = RolCoef_u,
          m_vehicle = `Mass (kg)`,
          grav_constant = Constants |>
            filter(Name == "grav_constant") |> pull(value),
          c_drag = `Aero_drag_coef (-)`,
          A_vehicle = `Surface_Area (m2)`,
          rho_air = Constants |>
            filter(Name == "rho_air") |> pull(value),
          v_start_decel = v_start_decel,
          v_end_decel = v_end_decel,
          v_wind = v_wind,
          alpha_slope = `Longitudinal slope (%)` / 100,
          m_rotate = m_rotate,
          c_decel = `Deceleration constant (m.s^-2)`),
        "Acceleration" =  f_accel_long_force(
          c_roll = RolCoef_u,
          m_vehicle = `Mass (kg)`,
          grav_constant = Constants |>
            filter(Name == "grav_constant") |> pull(value),
          c_drag = `Aero_drag_coef (-)`,
          A_vehicle = `Surface_Area (m2)`,
          rho_air = Constants |>
            filter(Name == "rho_air") |> pull(value),
          v_start_accel = v_start_accel,
          v_end_accel = v_end_accel,
          v_wind = v_wind,
          alpha_slope = `Longitudinal slope (%)` / 100,
          m_rotate = m_rotate,
          c_accel = `Acceleration constant (m.s^-2)`),
        "Constant" = f_const_speed_long_force(
          c_roll = RolCoef_u,
          m_vehicle = `Mass (kg)`,
          grav_constant = Constants |>
            filter(Name == "grav_constant") |> pull(value),
          c_drag = `Aero_drag_coef (-)`,
          A_vehicle = `Surface_Area (m2)`,
          rho_air = Constants |>
            filter(Name == "rho_air") |> pull(value),
          v_vehicle = v_const,
          v_wind = `v_wind`,
          alpha_slope = `Longitudinal slope (%)` /100)
      ),
    ForceLatt =
      switch(`Sector name`, 
             "East BendSIML" = 
               f_lat_force( m_vehicle = `Mass (kg)`,
                            grav_constant = Constants |> 
                              filter(Name == "grav_constant") |> pull(value),
                            r_corner = `Corner radius (m)`,
                            alpha_bank_slope = `Latitudinal slope (%)` / 100,
                            v_vehicle = v_const) ,
             "North StraightSIML" = 0
      )
  )

head(AllData)

```

## 3. Total Slip

```{r slip, echo=FALSE, message=FALSE, warning=FALSE}
# add slipt to the data:

AllData <-
  AllData |> 
  rowwise() |> 
  mutate(
    SlipLong =
      switch (`Maneuver code`,
              "Deceleration" =       
                f_decel_long_slip(c_roll = RolCoef_u, 
                                  m_vehicle = `Mass (kg)`, 
                                  grav_constant = Constants |> 
                                    filter(Name == "grav_constant") |> pull(value), 
                                  c_drag = `Aero_drag_coef (-)`, 
                                  A_vehicle = `Surface_Area (m2)`,
                                  rho_air = Constants |> 
                                    filter(Name == "rho_air") |> pull(value),
                                  v_start_decel = v_start_decel, 
                                  v_end_decel = v_end_decel, 
                                  v_wind = `v_wind`,
                                  alpha_slope = `Longitudinal slope (%)`/100, 
                                  m_rotate = m_rotate, 
                                  c_decel = `Deceleration constant (m.s^-2)`, 
                                  optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                                  grip_index_tyre = GripIndex_u,
                                  wet_mu_max_ref_tyre = Constants |>
                                    filter(Name == "mu_max_ref_tyre_wet") |> pull(value),
                                  x_correct_mu_max_track = x_correct_mu_max_track_u,
                                  c_brake_ref_tyre_wet = Constants |> 
                                    filter(Name == "c_brake_ref_tyre_wet") |> pull(value)),
              "Acceleration" =  
                f_accel_long_slip(c_roll = RolCoef_u, 
                                  m_vehicle = `Mass (kg)`, 
                                  grav_constant = Constants |> 
                                    filter(Name == "grav_constant") |> pull(value), 
                                  c_drag = `Aero_drag_coef (-)`, 
                                  A_vehicle = `Surface_Area (m2)`, 
                                  rho_air = Constants |> 
                                    filter(Name == "rho_air") |> pull(value),
                                  v_start_accel = v_start_accel, 
                                  v_end_accel = v_end_accel, 
                                  v_wind = v_wind, 
                                  alpha_slope = `Longitudinal slope (%)`/100, 
                                  m_rotate = m_rotate, 
                                  c_accel = `Acceleration constant (m.s^-2)`, 
                                  optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                                  grip_index_tyre = GripIndex_u, 
                                  wet_mu_max_ref_tyre = Constants |> 
                                    filter(Name == "mu_max_ref_tyre_wet") |> pull(value),
                                  x_correct_mu_max_track = x_correct_mu_max_track_u),
              "Constant" =       
                f_const_speed_long_slip(c_roll = RolCoef_u, 
                                        m_vehicle = `Mass (kg)`, 
                                        m_rotate = m_rotate,
                                        grav_constant = Constants |> 
                                          filter(Name == "grav_constant") |> pull(value), 
                                        c_drag = `Aero_drag_coef (-)`, 
                                        A_vehicle = `Surface_Area (m2)`, 
                                        rho_air = Constants |> 
                                          filter(Name == "rho_air") |> pull(value), 
                                        v_vehicle = v_const, 
                                        v_wind = `v_wind`,
                                        alpha_slope = `Longitudinal slope (%)`/100,
                                        wet_mu_max_ref_tyre = Constants |> 
                                          filter(Name == "mu_max_ref_tyre_wet") |> pull(value), 
                                        optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                                        grip_index_tyre = GripIndex_u, 
                                        c_brake_ref_tyre_wet = Constants |> 
                                          filter(Name == "c_brake_ref_tyre_wet") |> pull(value), 
                                        x_correct_mu_max_track = x_correct_mu_max_track_u)
      ),
    SlipLatt =
      switch(`Sector name`, 
             "East BendSIML" = 
               f_lat_slip(m_vehicle = `Mass (kg)`, 
                          v_vehicle = v_const, 
                          r_corner = `Corner radius (m)`, 
                          grav_constant = Constants |> 
                            filter(Name == "grav_constant") |> pull(value),
                          alpha_bank_slope = `Latitudinal slope (%)`/100, 
                          optimal_slip_ratio_tyre_track = optimal_slip_ratio_track_u,
                          grip_index_tyre = GripIndex_u, 
                          wet_mu_max_ref_tyre = Constants |> 
                            filter(Name == "mu_max_ref_tyre_wet") |> pull(value), 
                          x_correct_mu_max_track = x_correct_mu_max_track_u),
             "North StraightSIML" = 0
      )
  )

AllData

```

## 4. Total Friction Work

```{r FrictionWork, echo=FALSE, message=FALSE, warning=FALSE}

AllData_fw <- 
  AllData |> rowwise() |>  
  mutate(
    ManeuverDistance = switch(
      `Maneuver code`,
      "Deceleration" =     
        f_decel_distance(v_start = v_start_decel,
                         v_end = v_end_decel,
                         c_decel = `Deceleration constant (m.s^-2)`),
      "Acceleration" =
        f_accel_distance(v_start = v_start_accel ,
                         v_end = v_end_accel,
                         c_accel = `Acceleration constant (m.s^-2)`),
      "Constant" =  {switch(`Sector name`,
                            "North StraightSIML" = 
                              `Sector Distance (m)`-
                              f_decel_distance(v_start = v_start_decel,
                                               v_end = v_end_decel,
                                               c_decel = `Deceleration constant (m.s^-2)`)-
                              f_accel_distance(v_start = v_start_accel ,
                                               v_end = v_end_accel,
                                               c_accel = `Acceleration constant (m.s^-2)`),
                            "East BendSIML"  = 
                              `Sector Distance (m)`)
      }),
    FrictionWork_Man_Long = ForceLong*SlipLong*ManeuverDistance*`Maneuver repeats`,
    FrictionWork_Man_Latt = ForceLatt*SlipLatt*ManeuverDistance*`Maneuver repeats`
  )


# Get bad runs:
# AllData_fw |> filter(ManeuverDistance < 0) # Runs where acceleration and/or deceleration takes longer than total distance are an artifact so should be removed

bad_runs <- c(AllData_fw |> filter(ManeuverDistance < 0) |>  distinct(RUNS) |> pull(RUNS))
AllData_fw <-
  AllData_fw |> filter(!(RUNS %in% bad_runs))

## FW per sector (Sum over maneuver number)
# AllData_fw |> distinct(`Sector name`)

SectData_fw <- 
  AllData_fw |> ungroup() |> 
  group_by(RUNS, `Sector name`,`Test section`) |> 
  summarise(FrictionWork_sect_Long = sum(FrictionWork_Man_Long),
            FrictionWork_sect_Latt = sum(FrictionWork_Man_Latt),
            Distance_Sector = sum(ManeuverDistance)) |> 
  mutate(
    FrictionWork_sect_Long_pm = FrictionWork_sect_Long/Distance_Sector,
    FrictionWork_sect_Latt_pm = FrictionWork_sect_Latt/Distance_Sector,
    FrictionWork_sect_pm = (FrictionWork_sect_Long+FrictionWork_sect_Latt)/Distance_Sector
  )

## FW per test (sum over test number)
TrackData_fw <- 
  SectData_fw |> ungroup() |> 
  group_by(RUNS, `Test section`) |> 
  summarise(FrictionWork_Long = sum(FrictionWork_sect_Long),
            FrictionWork_Latt = sum(FrictionWork_sect_Latt),
            Distance_Track = sum(Distance_Sector)) |> 
  mutate(
    FrictionWork_Long_pm = FrictionWork_Long/Distance_Track,
    FrictionWork_Latt_pm = FrictionWork_Latt/Distance_Track,
    FrictionWork_pm = (FrictionWork_Long+FrictionWork_Latt)/Distance_Track,
    FrictionWork_Latt_7x_pm = 7*FrictionWork_Latt/Distance_Track,
    FrictionWork_7x_pm = (FrictionWork_Long+FrictionWork_Latt*7)/Distance_Track
  )


# inputs


Prob_X <- 
  PROB_X |> unnest(cols = where(is.list)) |> 
  filter(!(RUNS %in% bad_runs)) 

# select(-RUNS) |> 
#   mutate(across(everything(), ~log(.x+100)))

# Prob_Y <- TrackData_fw

# FrWrk_type = "FrictionWork_pm"
## sens per test

GSA_helper <- function(Prob_X, Prob_Y, FrWrk_type){
  
  probXY <- left_join(
    Prob_Y |> select(all_of(FrWrk_type)) |> mutate(across(all_of(FrWrk_type), ~ na_if(.x, 0)))
    # transform inputs by adding a constant and taking the logarithm
    ,  Prob_X |> mutate(across(-RUNS, ~log(.x+100))))  |> 
    # transform output taking logarithm
    mutate( across(all_of(FrWrk_type),~  log(.x))) |> 
    drop_na() |> ungroup() |> select(-RUNS)
  
  PROB_X= select(probXY,-any_of(FrWrk_type))
  PROB_Y=  pull(probXY,all_of(FrWrk_type))
  
  m1 <- sensiFdiv(model = NULL, X=PROB_X, fdiv = "TV", nboot = 0, conf = 0.95, scale = TRUE)
  tell(m1, y=PROB_Y, S)
  
  borg_df <- tibble(colnames(PROB_X),m1$S$original)
  names(borg_df)<-c("Factor", "delta")
  borg_df$Scenario<- FrWrk_type
  
  borg_df
  
}

TrackData_fw


```


## 5. Global Sensitivity Analysis
```{r  echo=FALSE, message=FALSE, warning=FALSE}
library(sensitivity)
library(ks)
# test
FrWrk_type_combi <- c("FrictionWork_Long_pm", "FrictionWork_Latt_pm", "FrictionWork_pm","FrictionWork_7x_pm")

GSA_output1 <- GSA_helper (Prob_X=Prob_X, Prob_Y=TrackData_fw, 
                           FrWrk_type=FrWrk_type_combi[1])

for (i in 2:length(FrWrk_type_combi)){
  GSA_output1 <- bind_rows(GSA_output1,
                           GSA_helper (Prob_X=Prob_X, Prob_Y=TrackData_fw, 
                                       FrWrk_type=FrWrk_type_combi[i]))
}

ggplot(GSA_output1, aes(x=Scenario, y=Factor, fill= delta)) + 
  geom_tile()+labs(x = "Type of friction work")+
  theme(axis.text.y = element_text(size = 8),axis.text.x = element_text(size = 8,angle = 45,hjust = 1)) 


```
